\section{Conclusion}
\label{sec:conclusion}

This paper presented the \emph{assurance passport}, a W3C verifiable credential profile that combines consortium blockchain with zero-knowledge proofs to provide privacy-preserving trust for Digital Mission Engineering.
We made three contributions.

First, we defined the assurance passport as a domain-specific VC profile for engineering artifact attestation, specifying four ZK proof predicates---schema compliance, version validity, authorized origin, and derivation lineage---under a multi-stakeholder threat model with five security goals.
The framework supports self-issued credentials with trust anchored in mathematical proof soundness and on-chain commitment immutability, complemented by endorsement credentials for governance approval.

Second, we presented a portable system architecture demonstrated on a GoQuorum consortium blockchain, Groth16 zk-SNARK circuits with Poseidon hashing, and a production DME platform (Kompozition) via a non-invasive adapter pattern.
The architecture uses a unified hash function (Poseidon) both on-chain and in-circuit, eliminating the Keccak/Poseidon equivalence problem, and localizes vendor-specific logic to the serialization layer.

Third, we deployed and evaluated a proof-of-concept on edge hardware (four NVIDIA Jetson Orin Nano boards), benchmarking proof generation, on-chain verification, and throughput across two BFT consensus protocols (IBFT, QBFT) and a CFT baseline (Raft).
The evaluation demonstrates that ZK-based artifact attestation is feasible on resource-constrained hardware, with proof generation as the performance bottleneck and on-chain verification remaining lightweight via BN254 pairing precompiles.

The assurance passport addresses a gap at the intersection of blockchain-based engineering trust, zero-knowledge provenance, and verifiable credentials: no prior work combines these three elements for engineering artifact attestation.
By enabling stakeholders to prove structural properties of artifacts without revealing proprietary content, the framework resolves the tension between transparency for trust and confidentiality for intellectual property protection---a fundamental barrier to artifact sharing in coalition DME environments.

Future work includes extending the predicate set to V\&V attestation, supporting larger artifacts via Merkle-based incremental proofs, exploring recursive proof composition for constant-time bundle verification, and validating the framework in a multi-organization coalition deployment.
