\section{Discussion}
\label{sec:discussion}

This section examines the limitations, design trade-offs, and broader implications of the assurance passport framework.

\subsection{Limitations}
\label{sec:discussion-limitations}

\textbf{Integrity without correctness.}
As noted in Section~\ref{sec:scope}, the assurance passport guarantees structural integrity and provenance but not the technical correctness of the underlying artifact.
A legitimately authored but flawed simulation model will receive a valid passport.
This ``garbage in, garbage out'' limitation is inherent to any integrity-focused system and is complementary to, not a replacement for, Verification \& Validation (V\&V) processes.

\textbf{Commitment hiding.}
The current commitment scheme uses a bare hash: $c_i^{v} = H(a_i^{v})$.
This provides binding (the committer cannot later claim a different artifact) but is not computationally hiding---an adversary who can guess the artifact content can verify their guess by hashing and comparing against the on-chain commitment.
For engineering artifacts with sufficient structural complexity and entropy (e.g., full SysML models), brute-force guessing is infeasible.
However, for small or predictable artifacts (e.g., a requirements document with a known template), a hiding commitment of the form $c_i^{v} = H(a_i^{v}, r)$, where $r$ is a random blinding factor included in the witness, would provide computational hiding.
We consider this an important production hardening step, requiring only a minor circuit modification (the blinding factor $r$ becomes part of the private witness).

\textbf{Concrete security level.}
BN254 provides approximately 100--110 bits of security following advances in the number field sieve for pairing-friendly curves.
While sufficient for this proof-of-concept, defense applications typically require 128-bit security.
Production deployments should migrate to BLS12-381, which provides the full 128-bit margin at the cost of larger proof sizes ($\sim$192 bytes compressed) and slower proof generation.

\textbf{Artifact size bounds.}
The current circuit design supports artifacts up to 256 field elements (approximately 8~KB serialized).
While this covers many individual DME work products (requirements packages, interface specifications, trade study summaries), large system-of-systems models can exceed this bound by orders of magnitude.
Scaling to larger artifacts requires either (a)~Merkle tree commitments over sub-artifact partitions, with each partition proved independently, or (b)~recursive proof composition (see Section~\ref{sec:discussion-future}).

\textbf{Deterministic transformation constraint.}
The derivation lineage predicate (Section~\ref{sec:proof-predicates}) requires that the transformation $f_\tau$ between parent and derived artifacts be deterministic.
Many real-world DME workflows involve non-deterministic steps (e.g., stochastic simulations, human-in-the-loop design refinement).
Extending the lineage predicate to probabilistic or approximate transformations is an open research problem.

\textbf{Serialization determinism.}
The commitment chain depends on deterministic serialization: the DME platform exports an artifact, the adapter serializes it to canonical JSON (RFC~8785), field elements are packed, and Poseidon hashes are computed.
RFC~8785 canonicalizes JSON key order and number/string representation, but does \emph{not} canonicalize the order of elements within JSON arrays.
SysML models contain semantically unordered collections (e.g., owned elements of a package) that are serialized as JSON arrays.
If the platform API returns these in non-deterministic order across calls, the canonical JSON diverges and the commitment changes.
The adapter must therefore impose a canonical element ordering within arrays (e.g., lexicographic by element identifier) beyond what RFC~8785 provides.
Cross-implementation consistency of the Poseidon hash (between the Circom in-circuit implementation and the Solidity on-chain library) must also be validated via shared test vectors to ensure that on-chain verification accepts in-circuit commitments.

\textbf{API projection scope.}
The commitment covers the artifact content as exported by the DME platform's REST API, which may be a projection of the internal model state (omitting tool-specific metadata, diagram layout, or internal identifiers).
The assurance passport therefore attests to properties of the API-visible artifact, not necessarily the complete internal model.
For the proof-of-concept, this distinction is acceptable; production deployments should document which model facets the API export captures.

\subsection{Design Trade-Offs}
\label{sec:discussion-tradeoffs}

\textbf{Groth16 and trusted setup.}
We selected Groth16 for its constant-size proofs (128~bytes compressed on BN254), constant-time verification (a single multi-pairing check over 4 input pairs), and mature tooling (Circom, SnarkJS).
The principal drawback is the per-circuit trusted setup: each predicate circuit requires a separate setup ceremony, and a compromised common reference string breaks soundness for all proofs generated with that circuit.
For production deployments, we recommend a multi-party computation (MPC) ceremony with at least one party per consortium stakeholder, following the protocol used in Zcash Sapling.
Alternatively, universal-setup systems such as PLONK or transparent systems such as STARKs eliminate the per-circuit trusted setup at the cost of larger proofs (PLONK: $\sim$400--800~bytes; STARKs: $\sim$40--200~KB).
For bandwidth-constrained DDIL environments, Groth16's proof compactness remains advantageous.

\textbf{GoQuorum deprecation.}
GoQuorum was selected for the proof-of-concept due to its mature Raft support, full EVM compatibility, and existing deployment on our edge hardware testbed.
ConsenSys has since shifted focus to Hyperledger Besu as the primary enterprise Ethereum client.
However, the assurance passport framework is not GoQuorum-specific: the smart contracts (Solidity), ZK verification (BN254 precompiles), and VC format are EVM-standard and portable to Besu, Polygon Edge, or any Ethereum-compatible consortium chain.
Migration requires redeploying contracts and reconfiguring consensus; no architectural changes are needed.

\textbf{Poseidon vs.\ Keccak.}
The choice of Poseidon over Keccak-256 for artifact commitments is driven by in-circuit efficiency ($\sim$300 vs.\ $\sim$150{,}000 R1CS constraints per hash invocation).
Using a single hash function for both on-chain and in-circuit computation eliminates the need for cross-hash equivalence proofs but introduces a non-standard on-chain hash.
Since GoQuorum is permissioned, consortium operators can deploy a Poseidon library contract (or native precompile) without impacting the broader Ethereum ecosystem.
This trade-off is appropriate for consortium chains but would require additional justification for public-chain deployments.

\textbf{Self-issuance vs.\ third-party attestation.}
The assurance passport uses a self-issuance model where the stakeholder who produces the artifact is also the credential issuer.
Trust is anchored not in the issuer's identity but in the combination of ZK proof soundness, on-chain commitment immutability, and the authorized origin predicate $\pi_{\text{origin}}$.
For programmatic acceptance decisions (e.g., milestone reviews), a separate endorsement credential issued by a designated authority provides the governance layer.
This two-layer separation (cryptographic assurance + governance approval) mirrors established defense acquisition practice, where engineering evidence and programmatic approval are distinct activities.

\subsection{DDIL Operational Considerations}
\label{sec:discussion-ddil}

Deploying the assurance passport in DDIL environments raises three operational concerns beyond performance.

\textbf{Consensus liveness.}
With $n=4$ validators under IBFT or QBFT, the system requires 3 of 4 nodes online for liveness.
In DDIL scenarios with intermittent connectivity, losing two nodes halts block production.
Mitigation strategies include:
\begin{itemize}
    \item Increasing the validator set (e.g., $n=7$, tolerating 2 faults).
    \item Queuing commitment transactions locally during network partitions and replaying them upon reconnection.
    \item Using Raft (crash fault tolerant, no BFT overhead) within a single-organization enclave where Byzantine behavior is not a concern, reserving IBFT/QBFT for cross-organizational boundaries.
\end{itemize}

\textbf{Clock synchronization.}
Block timestamps in the commitment registry provide temporal ordering.
DDIL environments may lack reliable NTP; however, IBFT/QBFT consensus provides a logical clock through block height, which is sufficient for ordering guarantees.
Absolute wall-clock accuracy is relevant only for expiration checks in the credential's validity window, which can be set conservatively.

\textbf{Network partitions.}
If the consortium network splits, each partition operates independently.
The BFT partition (with $\geq 3$ of 4 nodes) continues to finalize blocks; the minority partition halts.
Commitments made on the majority partition are authoritative.
A merge protocol for reconciling divergent state after partition healing is out of scope but is a relevant concern for extended DDIL operations.

\textbf{Resource contention.}
Each Jetson Nano serves simultaneously as a consortium validator and a proof generation endpoint.
Proof generation is CPU-intensive and can take minutes for a full bundle; during this period, the validator process must continue to participate in consensus rounds.
With $n=4$ and $f=1$, resource starvation of a single validator could threaten liveness.
OS-level resource isolation (e.g., cgroups, CPU pinning) should be used in production deployments to ensure the validator process is not preempted by the prover.
Our benchmarks measure proof generation and consensus performance separately; concurrent performance under load is a topic for future evaluation.

\textbf{Commitment-proof temporal gap.}
The assurance passport lifecycle commits the artifact hash on-chain (Phase~1) before proof generation begins (Phase~2).
Since proof generation takes minutes on edge hardware, a temporal window exists during which the commitment is anchored but no proofs are available.
The passport is not valid until all proofs are verified on-chain; verifiers must check both the commitment timestamp and the proof verification timestamp before accepting a passport.

\subsection{Implications for Coalition DME}
\label{sec:discussion-coalition}

The assurance passport framework has particular relevance for coalition DME environments, where multiple organizations (potentially from different nations) contribute artifacts to a shared ASOT under strict IP protection requirements.
The predicate attestation property (Security Goal~5) enables a stakeholder to prove that an artifact satisfies required properties---schema compliance, authorized provenance, version validity---without revealing the artifact content to coalition partners.
This addresses a fundamental barrier to MBSE artifact sharing in defense contexts: the tension between transparency for trust and confidentiality for IP protection.

The framework does not, however, address semantic interoperability across coalition toolchains.
Each coalition partner may use different SysML profiles, metamodels, or domain-specific extensions.
The canonical JSON serialization (Section~\ref{sec:arch-dme}) must be defined by consortium agreement.
Cross-platform commitment interoperability---ensuring that the same logical model produces the same commitment regardless of source platform---requires a shared canonical metamodel specification, a prerequisite for multi-vendor coalition deployments.
Achieving consensus on a canonical metamodel across competing defense contractors is a significant organizational challenge; ongoing SysML~v2 API standardization efforts may provide a foundation for future interoperability.

The framework operates within a single classification domain; cross-classification sharing (e.g., CUI to Secret) requires cross-domain solutions that are orthogonal to this work.
Similarly, ITAR/EAR export control constraints and DFARS data rights (252.227-7013/7014) shape what can be shared in coalition environments; the assurance passport addresses the cryptographic layer but does not replace contractual or regulatory governance.

The assurance passport can be viewed as a cryptographic layer over the \emph{digital thread}---the DoD's concept for end-to-end traceability of engineering data across the acquisition lifecycle---providing verifiable provenance at each link in the data chain.

\subsection{Future Work}
\label{sec:discussion-future}

\textbf{V\&V attestation predicate.}
A fifth proof predicate attesting to Verification \& Validation status would extend the passport's scope from structural integrity to programmatic acceptance.
This could be modeled as a ZK proof that a designated IV\&V agent's signature endorses a specific predicate outcome, or as a separate endorsement credential within the VC framework.

\textbf{Recursive proof composition.}
Replacing the current concatenation-based proof bundling with recursive proof composition (e.g., SnarkPack or Halo-style folding) would reduce verification cost from $O(k)$ to $O(1)$ pairing operations for $k$ predicates.
This is particularly attractive as the number of predicates grows beyond four.

\textbf{Large artifact support.}
Merkle tree commitments over artifact partitions, combined with incremental proof updates (proving only the changed sub-tree), would extend the framework to arbitrarily large engineering artifacts.
Incremental proving also amortizes proof generation cost across version updates.

\textbf{Formal circuit verification.}
The correctness of ZK circuits is a critical trust assumption.
Formal verification of the circuit implementations (e.g., using tools such as Ecne or Picus) would provide stronger assurance that the proof predicates correctly encode the intended properties.

\textbf{Multi-chain federation.}
For coalition environments spanning multiple consortium blockchains (e.g., one per national domain), a cross-chain bridge for commitment verification would enable federated ASOT operation without requiring all parties to share a single chain.
