\section{Introduction}
\label{sec:introduction}

Digital Mission Engineering (DME) has emerged as a critical paradigm for defense and aerospace organizations seeking to model, simulate, and evaluate mission outcomes before committing physical resources~\cite{dod_de_strategy}.
In a typical DME ecosystem, multiple stakeholders---prime contractors, subcontractors, government agencies, and academic partners---contribute engineering artifacts such as SysML models, simulation configurations, requirements specifications, and analysis results to a shared \emph{Authoritative Source of Truth} (ASOT).
The integrity of this ASOT is foundational: mission-critical decisions depend on the assurance that artifacts are authentic, unaltered, and traceable to their origin.

While modern DME platforms (e.g., Teamwork Cloud, Kompozition) provide element-level change tracking within their toolchains, current ASOT implementations lack \emph{cryptographic} tamper evidence, offer no immutable cross-organizational accountability trail when missions fail, and provide no mechanism for stakeholders to verify artifact provenance without full access to the underlying data~\cite{simba_chain}.
These environments must also operate under Denied, Degraded, Intermittent, and Limited (DDIL) connectivity conditions, where reliance on cloud-hosted infrastructure is not guaranteed.
This creates two compounding problems.
First, \emph{trust deficit}: when a mission outcome diverges from predictions, there is no verifiable record of what artifact versions existed at the time of a decision, who contributed them, or whether they were modified after the fact.
Second, \emph{intellectual property (IP) exposure}: stakeholders operating in coalition or cross-organizational settings are reluctant to share proprietary models openly, yet current approaches require full data visibility to establish trust.

Blockchain-based solutions have begun to address the first problem.
SIMBA Chain has deployed blockchain infrastructure for Department of Defense supply chain integrity~\cite{simba_chain}, and Lu et al.~\cite{lu2020decentralized} proposed a decentralized marketplace for Model-Based Systems Engineering (MBSE) artifacts.
These approaches provide immutability and decentralized consensus, and while consortium platforms such as GoQuorum support private transactions, they do not enable stakeholders to prove \emph{properties} of hidden data---a significant barrier in IP-sensitive environments where selective verification, not just access control, is required.
Meanwhile, zero-knowledge proof (ZKP) systems have demonstrated the ability to attest to computational properties without revealing underlying data, with applications in supply chain provenance~\cite{zpie}, financial compliance, and verifiable computation.
However, ZKP systems alone lack the immutability and decentralized consensus that blockchain provides.
To the best of our knowledge, no existing work combines ZKPs with blockchain for privacy-preserving provenance of \emph{engineering} artifacts---a domain where the data structures are significantly more complex than flat byte sequences or simple transactions, involving typed multi-relational graphs and cross-layer semantic dependencies.

In this paper, we propose the \emph{Assurance Passport}, a verifiable credential (VC) profile that combines consortium blockchain with zero-knowledge proofs to establish a vendor-neutral, privacy-preserving trust layer for DME.
An assurance passport is a W3C-aligned credential~\cite{w3c_vc} that attests to properties of an engineering artifact---schema compliance, version validity, authorized provenance, and derivation lineage---without disclosing the artifact's proprietary content.
By anchoring these credentials on a consortium blockchain, we provide immutable temporal integrity (``what existed when'') while ZK proofs enable zero-knowledge predicate attestation (``this artifact satisfies property $P$ without revealing the artifact itself'').

This work extends our prior investigation of on-chain ZK verification for AI model marketplaces~\cite{jagannath2025zksnark}, which established the feasibility of zk-SNARK verification on public blockchains but was limited to AI model performance claims on a public testnet.
The present work differs in three key dimensions: (1)~we target \emph{engineering artifacts} rather than AI models, requiring richer proof predicates over structured data; (2)~we deploy on a \emph{consortium blockchain} (GoQuorum) rather than a public testnet, reflecting the permissioned trust model of DME; and (3)~we evaluate on \emph{edge hardware} (NVIDIA Jetson) to address Denied, Degraded, Intermittent, and Limited (DDIL) operational environments where cloud connectivity cannot be assumed.

Our contributions are as follows:
\begin{enumerate}
    \item We define the \emph{assurance passport} as a W3C verifiable credential profile for engineering artifact attestation, specifying its schema, proof predicates, and security properties under a multi-stakeholder threat model.
    \item We present a vendor-neutral system architecture integrating GoQuorum consortium blockchain, ZK proof generation, and a production DME platform (Kompozition) via API.
    \item We evaluate a proof-of-concept deployment on edge hardware (four NVIDIA Jetson boards), benchmarking proof generation time, verification latency, gas costs, and throughput across three consensus protocols (IBFT, QBFT, Raft).
\end{enumerate}

The remainder of this paper is organized as follows.
Section~\ref{sec:related-work} surveys related work in blockchain-based trust for engineering, ZK provenance systems, and DME frameworks.
Section~\ref{sec:preliminaries} provides necessary background on ZK-SNARKs, consortium blockchains, and verifiable credentials.
Section~\ref{sec:threat-model} formalizes our threat model.
Section~\ref{sec:framework} defines the assurance passport framework.
Section~\ref{sec:architecture} describes the system architecture.
Section~\ref{sec:evaluation} presents experimental results.
Section~\ref{sec:discussion} discusses limitations and implications, and Section~\ref{sec:conclusion} concludes.
