\section{Related Work}
\label{sec:related-work}

We organize related work along three axes: blockchain for engineering trust (Section~\ref{sec:rw-blockchain-eng}), zero-knowledge proofs for provenance and privacy (Section~\ref{sec:rw-zk-provenance}), and verifiable credential profiles for assurance (Section~\ref{sec:rw-vc-profiles}).
Table~\ref{tab:related-work} summarizes how existing approaches compare to the assurance passport.

\subsection{Blockchain for Engineering Trust}
\label{sec:rw-blockchain-eng}

The application of blockchain to systems engineering has progressed from conceptual proposals to funded prototypes.
Kuhn et al.~\cite{kuhn2017digital_twin_blockchain} were among the first to propose blockchain for traceability in the context of digital twins and MBSE, focusing on electrical/electronic (E/E) architecture artifacts.
Suhail et al.~\cite{suhail2022bcdt} provide a comprehensive survey of blockchain-based digital twins, identifying provenance-enabled lineage tracking as a key capability but noting that no existing work applies zero-knowledge proofs to digital twin provenance.
Yue, Guerra, and Wagner~\cite{yue2021blockchain_mbse} conducted a pilot blockchain experiment on an aerospace MBSE project using Hyperledger Fabric, identifying governance complexity, compliance requirements, and privacy rules as key challenges for consortium deployments.
In a follow-up study, Yue and Guerra~\cite{yue2023confidentiality} directly addressed confidentiality and data integrity in consortium blockchain for MBSE, proposing access control mechanisms to protect intellectual property between participating organizations.
This work is the closest to ours on the blockchain--engineering axis; however, their approach relies on permissioned access control rather than cryptographic privacy, meaning that authorized participants still have full visibility into shared data.

Lu et al.~\cite{lu2020decentralized} proposed a decentralized marketplace for MBSE artifacts using distributed ledger technology, enabling cross-organizational exchange of engineering models.
Their architecture addresses decentralization but does not provide mechanisms for privacy-preserving verification---stakeholders must trust the marketplace with full artifact content.
Harz and Bader~\cite{harz2020permissioned} provided a systematic analysis of permissioned blockchains in the systems engineering lifecycle, arguing that consortium arrangements are most suitable for the SE context due to governance and compliance requirements.

On the industry side, SIMBA Chain has deployed blockchain infrastructure under U.S. Department of Defense contracts, including the ALAMEDA project (a ``single source of truth'' on distributed ledger for the U.S. Navy and Defense Logistics Agency) and the BASECAMP project for additive manufacturing traceability with the U.S. Air Force~\cite{simba_chain}.
These represent the largest real-world deployments of blockchain for defense engineering, but they focus on physical supply chain integrity rather than model-level provenance, and do not incorporate zero-knowledge proofs.

Beyond blockchain-based approaches, the open-source OpenMBEE platform (developed by JPL/NASA) provides version-controlled ASOT capabilities for MBSE, and the OSLC (Open Services for Lifecycle Collaboration) standard enables cross-tool integration.
These solutions address governance and interoperability but lack cryptographic tamper evidence or privacy-preserving verification.

In terms of consensus performance, Wells et al.~\cite{wells2023hotstuff_goquorum} benchmarked IBFT, QBFT, and Basic HotStuff in private GoQuorum networks under varying network sizes and conditions, finding that Basic HotStuff outperformed IBFT and QBFT under low-bandwidth conditions.
However, their evaluation did not include ZK proof verification overhead, leaving open the question of how cryptographic computation interacts with consensus on resource-constrained nodes---a gap our evaluation addresses.

\subsection{Zero-Knowledge Proofs for Provenance and Privacy}
\label{sec:rw-zk-provenance}

Zero-knowledge proof systems have seen rapid adoption for privacy-preserving verification across domains.
Groth~\cite{groth2016size} established the foundational zk-SNARK construction (Groth16) that remains widely deployed due to its constant-size proofs and fast verification.
In supply chain contexts, ZKPs have been applied to verify trade compliance, document authenticity, and ethical sourcing claims without disclosing confidential business data~\cite{zkp_supply_chain2024, maritime_zkp2025}.

Two works are particularly relevant to our contribution.
First, Salleras and Daza~\cite{salleras2021zpie} developed ZPiE, a C library for executing zk-SNARK proofs on embedded systems, demonstrating Groth16 proof generation on a Raspberry Pi Zero W (700~MHz ARM).
Their results establish that ZK computation on resource-constrained hardware is feasible, though they did not integrate with a blockchain layer or target engineering workloads.
Second, Ramezan and Meamari~\cite{ramezan2024zkiot} proposed zk-IoT, a framework combining ZKPs with blockchain to ensure firmware integrity on IoT devices.
zk-IoT is the closest technical analog to our work on the ZK--blockchain--device axis, but it operates on firmware binary blobs rather than structured engineering artifacts, and addresses device authentication rather than artifact provenance.

Our prior work~\cite{jagannath2025zksnark} established the feasibility of on-chain zk-SNARK verification for AI model marketplaces, achieving 233.63\,s proof generation and 61.50\,s verification on the Sepolia public testnet.
The present work extends this in three dimensions: from AI models to engineering artifacts (requiring richer proof predicates), from public testnet to consortium blockchain (reflecting DME trust models), and from cloud to edge hardware (addressing DDIL environments).

\subsection{Verifiable Credential Profiles for Assurance}
\label{sec:rw-vc-profiles}

The W3C Verifiable Credentials Data Model v2.0~\cite{w3c_vc} provides a standard for expressing and verifying digital credentials.
Domain-specific VC profiles have emerged for several sectors: Open Badges 3.0~\cite{open_badges3} aligns with W3C VCs for educational achievement credentials, and the UN Transparency Protocol (UNTP)~\cite{untp_dpp} defines VC-based Digital Product Passports for supply chain sustainability.
The EU Battery Regulation~\cite{eu_battery_passport} will mandate VC-based product passports with conformity data and tiered stakeholder access from 2027---the first legally mandated sector-specific VC profile at scale.
Architectural guidance for implementing Digital Product Passports using decentralized identifiers (DIDs) and VCs with selective disclosure has been provided by recent work~\cite{dpp_did_vc2024}.
On the cryptographic side, BBS+ signatures~\cite{tessaro2023bbs} enable attribute-level selective disclosure natively within VC presentations---a holder can reveal a subset of credential fields without a full ZK proof.
However, BBS+ supports only field-level reveal (``show attribute $A$ but not $B$''), not predicate attestation over hidden data (``prove property $P(x)$ without revealing $x$'').
The assurance passport requires the latter: proving that a hidden artifact satisfies structural and relational predicates, which necessitates general-purpose ZK circuits rather than signature-based selective disclosure.

Despite this momentum, no VC profile exists for engineering assurance or DME artifacts.
The assurance passport proposed in this paper fills this gap by defining a VC schema for engineering artifact attestation, analogous to how Open Badges serves education and UNTP serves sustainability.

\subsection{Summary and Positioning}
\label{sec:rw-summary}

\begin{table*}[t]
\centering
\caption{Comparison of related approaches. \checkmark = fully addressed, $\sim$ = partially addressed, --- = not addressed.}
\label{tab:related-work}
\begin{tabular}{lccccc}
\toprule
\textbf{Approach} & \textbf{Blockchain} & \textbf{ZK Privacy} & \textbf{Engineering} & \textbf{Edge HW} & \textbf{VC Profile} \\
\midrule
Yue \& Guerra~\cite{yue2021blockchain_mbse, yue2023confidentiality} & \checkmark & --- & \checkmark & --- & --- \\
Lu et al.~\cite{lu2020decentralized} & \checkmark & --- & \checkmark & --- & --- \\
SIMBA Chain~\cite{simba_chain} & \checkmark & --- & $\sim$ & --- & --- \\
ZPiE~\cite{salleras2021zpie} & --- & \checkmark & --- & \checkmark & --- \\
zk-IoT~\cite{ramezan2024zkiot} & \checkmark & \checkmark & --- & $\sim$ & --- \\
Jagannath et al.~\cite{jagannath2025zksnark} & \checkmark & \checkmark & --- & --- & --- \\
UNTP/DPP~\cite{untp_dpp} & --- & $\sim$ & --- & --- & \checkmark \\
EU Battery Passport~\cite{eu_battery_passport} & --- & $\sim$ & --- & --- & \checkmark \\
BBS+ Signatures~\cite{tessaro2023bbs} & --- & $\sim$ & --- & --- & $\sim$ \\
\midrule
\textbf{Assurance Passport (ours)} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table*}

As Table~\ref{tab:related-work} illustrates, existing work addresses subsets of the design space but no prior approach combines all five dimensions.
Blockchain-based engineering trust systems (Yue \& Guerra, Lu et al., SIMBA Chain) lack ZK privacy.
ZK systems on constrained hardware (ZPiE) lack blockchain integration.
ZK--blockchain systems (zk-IoT, our prior work) do not target engineering artifacts.
VC profiles (UNTP, EU Battery Passport) exist for sustainability and product compliance but not for engineering assurance.
BBS+ signatures provide attribute-level selective disclosure but not predicate attestation over hidden structured data.
The assurance passport is the first framework, to our knowledge, to unify consortium blockchain, zero-knowledge proofs, engineering artifact provenance, edge deployment, and a W3C-aligned VC profile.

A natural question is whether ZK proofs add meaningful value over simple signed hashes stored on-chain.
Signed hashes prove that a known party committed to specific content, but require the verifier to access the full artifact to check any property.
ZK proofs enable \emph{property verification without content disclosure}: a verifier can confirm schema compliance, version validity, authorized origin, or derivation lineage without seeing the artifact itself.
This distinction is critical in coalition environments where IP protection prevents full data sharing.
