\section{Threat Model}
\label{sec:threat-model}

We consider a DME ecosystem in which $n$ stakeholders $\{S_1, S_2, \ldots, S_n\}$ contribute engineering artifacts to a shared ASOT.
Each stakeholder controls one or more artifacts and may use different toolchains (e.g., SysML modeling tools, simulation platforms, requirements management systems).
A \emph{mission decision} $d$ is taken at time $t$ based on the set of artifact versions $\{a_1^{v_1}, a_2^{v_2}, \ldots, a_k^{v_k}\}$ present in the ASOT at that time.
We define the following adversarial roles, security goals, and explicit scope limitations.

\subsection{Adversaries}
\label{sec:adversaries}

\textbf{A1: Malicious Stakeholder.}
A stakeholder $S_i$ deliberately submits a falsified artifact $a_i' \neq a_i$ or retroactively modifies a previously committed artifact to alter the decision basis.
The adversary's goal is to influence mission outcomes or shift blame post-failure without detection.

\textbf{A2: Negligent Stakeholder.}
A stakeholder $S_j$ updates an artifact $a_j^{v} \rightarrow a_j^{v+1}$ without notifying dependent stakeholders or obtaining required approvals.
Unlike A1, the modification is not deliberately malicious but violates the change management protocol, potentially invalidating downstream analyses that depend on $a_j^{v}$.

\textbf{A3: Unauthorized Contributor.}
An entity not authorized to contribute to artifact $a_k$ submits a version, or an authorized stakeholder uses an unapproved tool $T' \notin \mathcal{T}_{\text{approved}}$ to generate the artifact.
This captures both external intrusion and insider misuse of unapproved toolchains.

\textbf{A4: Curious Observer.}
A stakeholder $S_m$ with legitimate access to the ASOT attempts to extract proprietary content from artifacts contributed by other stakeholders.
In coalition environments, this represents the IP exfiltration risk that prevents open sharing.

\subsection{Security Goals}
\label{sec:security-goals}

The assurance passport framework aims to provide the following properties:

\begin{enumerate}
    \item \textbf{Tamper Evidence (against A1).} Any modification to a committed artifact after time $t$ is detectable. Formally, given a commitment $c_i^{v} = H(a_i^{v})$ anchored on-chain at time $t$, any attempt to substitute $a_i' \neq a_i^{v}$ is detected with overwhelming probability (by collision resistance of $H$).

    \item \textbf{Temporal Integrity (against A1, A2).} For any decision $d$ taken at time $t$, the exact set of artifact versions $\{a_1^{v_1}, \ldots, a_k^{v_k}\}$ that constituted the ASOT at time $t$ can be reconstructed and verified. No stakeholder can retroactively claim a different version was current, as the on-chain record is immutable.

    \item \textbf{Change Accountability (against A2).} Every artifact version transition $a_j^{v} \rightarrow a_j^{v+1}$ is attributable to a specific stakeholder $S_j$ with a verifiable timestamp, enabling audit trails for change management compliance.

    \item \textbf{Authorized Provenance (against A3).} Each artifact version carries a verifiable attestation that it was produced by an authorized stakeholder $S_i \in \mathcal{S}_{\text{auth}}(a)$ using an approved tool $T \in \mathcal{T}_{\text{approved}}$. This is verified without revealing the artifact content.

    \item \textbf{Predicate Attestation (against A4).} A stakeholder can prove that an artifact satisfies a property $P$ (e.g., schema compliance, derivation from a specific parent artifact) without revealing the artifact itself. Formally, there exists a zero-knowledge proof $\pi$ such that a verifier is convinced that $P(a_i^{v}) = \texttt{true}$ while learning nothing about $a_i^{v}$ beyond the truth of $P$ and the public commitment $c_i^{v} = H(a_i^{v})$. Note that the commitment is binding but not hiding (see Section~\ref{sec:discussion-limitations}); an adversary who can guess the artifact content can verify the guess against the on-chain commitment.
\end{enumerate}

\subsection{Proof Predicates}
\label{sec:proof-predicates}

To distinguish the assurance passport from simple signed hashes, we define four classes of ZK proof predicates that go beyond hash-correctness.
Each predicate is realized as a zk-SNARK circuit with a public input $x$ (visible to verifiers) and a private witness $w$ (the artifact content, hidden from verifiers).
Given a common reference string from a trusted setup ceremony (see Section~\ref{sec:scope} for trust implications), a prover generates $\pi \leftarrow \textsf{Prove}(\mathsf{pk}, x, w)$ and a verifier checks $\textsf{Verify}(\mathsf{vk}, x, \pi) \rightarrow \{0, 1\}$.

\begin{enumerate}
    \item \textbf{Schema Compliance.} The artifact conforms to a restricted structural schema $S$, encompassing attribute presence checks, type assertions, and bounded-depth graph traversals over the serialized representation (see Section~\ref{sec:arch-zk} for the precise schema compliance language; full XSD or JSON Schema validation is not arithmetized).
    Public input: on-chain commitment $c_i^{v} = H(a_i^{v})$ and schema identifier $S$. Witness: artifact content $a_i^{v}$.
    \begin{equation}
    \label{eq:schema}
    \pi_{\text{schema}} \leftarrow \textsf{Prove}\!\left(\mathsf{pk},\; (c_i^{v}, S),\; a_i^{v}\right) \quad \text{s.t.} \quad H(a_i^{v}) = c_i^{v} \;\wedge\; a_i^{v} \models S
    \end{equation}

    \item \textbf{Version Validity.} The artifact is a valid update of its predecessor according to update rules $R$.
    $R$ is a consortium-configurable set of structural constraints---for example, no deletion of required elements, monotonic version counter, and preservation of interface signatures.
    Public input: commitments $c_i^{v}, c_i^{v+1}$. Witness: both artifact versions and the delta $\Delta$.
    \begin{equation}
    \label{eq:version}
    \pi_{\text{version}} \leftarrow \textsf{Prove}\!\left(\mathsf{pk},\; (c_i^{v}, c_i^{v+1}),\; (a_i^{v}, a_i^{v+1}, \Delta)\right)
    \end{equation}

    \item \textbf{Authorized Origin.} The artifact was produced by an authorized stakeholder using an approved tool.
    Public input: commitment $c_i^{v}$, Merkle roots of $\mathcal{S}_{\text{auth}}(a)$ and $\mathcal{T}_{\text{approved}}$. Witness: stakeholder credential, tool identity, Merkle paths.
    \begin{equation}
    \label{eq:origin}
    \pi_{\text{origin}} \leftarrow \textsf{Prove}\!\left(\mathsf{pk},\; (c_i^{v}, \mathsf{root}_S, \mathsf{root}_T),\; (S_i, T, \mathsf{path}_S, \mathsf{path}_T)\right)
    \end{equation}

    \item \textbf{Derivation Lineage.} The artifact was derived from a specific parent artifact using a known deterministic transformation $f_\tau$.
    Public input: commitments $c_i^{v}, c_j^{v'}$ and transformation identifier $\tau$. Witness: both artifact contents.
    \begin{equation}
    \label{eq:lineage}
    \pi_{\text{lineage}} \leftarrow \textsf{Prove}\!\left(\mathsf{pk},\; (c_i^{v}, c_j^{v'}, \tau),\; (a_i^{v}, a_j^{v'})\right)
    \end{equation}
\end{enumerate}

These predicates are bundled into an assurance passport credential, as defined in Section~\ref{sec:framework}.

\subsection{Scope and Limitations}
\label{sec:scope}

\textbf{What we do not cover.}
The assurance passport guarantees \emph{integrity}, \emph{provenance}, and \emph{predicate attestation}---it does not guarantee the \emph{correctness} of the source data.
If a stakeholder submits a legitimately authored but technically flawed simulation model, the system will faithfully record its provenance and attest to its properties, but will not detect the engineering error.
This ``garbage in, garbage out'' (GIGO) limitation is inherent to any integrity-focused system and is explicitly out of scope.

\textbf{Trust assumptions.}
We assume that (1)~the consortium blockchain operates under an honest-supermajority assumption among validator nodes (i.e., fewer than one-third are Byzantine, as required by IBFT/QBFT consensus), (2)~the ZK proof system is knowledge-sound under the Knowledge-of-Exponent assumption (KEA), which is non-falsifiable but standard for all deployed pairing-based SNARKs, (3)~the trusted setup ceremony was performed correctly (i.e., toxic waste was destroyed), (4)~each stakeholder's signing key is not compromised, and (5)~the ZK circuit implementations correctly encode the intended proof predicates (a buggy circuit could produce valid proofs for non-compliant artifacts; formal circuit verification is discussed in Section~\ref{sec:discussion-future}).
In a multi-stakeholder consortium, the trusted setup is a non-trivial trust point: a compromised common reference string would break soundness for all proof predicates.
We recommend a multi-party computation (MPC) ceremony, as used in Zcash and other deployed systems, to distribute trust across participants.
Key management and revocation are orthogonal concerns addressed by standard PKI infrastructure.

\textbf{Collusion bound.}
The system tolerates up to $\lfloor (n-1)/3 \rfloor$ colluding Byzantine validators under IBFT/QBFT.
Collusion among \emph{stakeholders} (as opposed to validators) is partially addressed: colluding stakeholders cannot forge provenance for artifacts they did not produce, but they can coordinate to submit complementary false artifacts that individually pass all proof predicates.
Detecting semantically coordinated falsification across artifacts remains an open problem.
